Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$. We prove in Appendix~\ref{app:hv}:

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take any $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i] = \bigcap_{i=1}^{k-2} r'_i$;
    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
\myeqn{
    \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}} \label{eqn:dcross-H} \\
    \dcross_V(\bm{t}) = \set{v \in V \mid \text{$v$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}}. \nn
    %\label{eqn:dcross-H} 
}
The prefix ``d-'' stands for ``double''. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:

@@

Next, we construct from $R_i'$ a new set of rectangles: 
\myeqn{
    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
}
In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.

@@

Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.

@@

Now, solve a $(k-1)$-SJ problem on the input $\set{R_1, ..., R_{k-3}, H', V}$ using the algorithm $\A$ supplied (by Lemma~\ref{lmm:hv}). Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ, whose input size is at most $n$. Given the lemma below (which is proved in Appendix~\ref{app:hv}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.

@@

We prove the next lemma in Appendix~\ref{app:hv} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).

@@

Next, we construct a new set of horizontal segments: 
\myeqn{
    H^* = \set{h^* \mid h' \in H' \text{ and its minimal segment $h^*$ exists}}. \label{eqn:hv:type2:H*}
}
This can be done in $O(n + k \cdot \out)$ time, as shown in Appendix~\ref{app:hv}.

@@

The proof can be found in Appendix~\ref{app:hv}.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} \\
\underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$. Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$. As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$. Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$. It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\
\underline{\em Proof of Statement (1).} We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t}) = \cross_H(r^*)$. After this, we can then derive $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$, where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.

>>>>

We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$. We prove in Appendix~\ref{app:hv:type1}:

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take any $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i] = \bigcap_{i=1}^{k-2} r'_i$;
    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
\myeqn{
    \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}} \label{eqn:dcross-H} \\
    \dcross_V(\bm{t}) = \set{v \in V \mid \text{$v$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}}. \nn
    %\label{eqn:dcross-H} 
}
The prefix ``d-'' stands for ``double''. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv:type1}:

@@

Next, we construct from $R_i'$ a new set of rectangles: 
\myeqn{
    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
}
In Appendix~\ref{app:hv:type1}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.

@@

Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv:type1}, presents some useful properties of these sets.

@@

Now, solve a $(k-1)$-SJ problem on the input $\set{R_1, ..., R_{k-3}, H', V}$ using the algorithm $\A$ supplied (by Lemma~\ref{lmm:hv}). Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ, whose input size is at most $n$. Given the lemma below (which is proved in Appendix~\ref{app:hv:type2}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.

@@

We prove the next lemma in Appendix~\ref{app:hv:type2} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).

@@

Next, we construct a new set of horizontal segments: 
\myeqn{
    H^* = \set{h^* \mid h' \in H' \text{ and its minimal segment $h^*$ exists}}. \label{eqn:hv:type2:H*}
}
This can be done in $O(n + k \cdot \out)$ time, as shown in Appendix~\ref{app:hv:type2}.

@@

The proof can be found in Appendix~\ref{app:hv:type2}.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
\underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$. Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$. As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$. Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$. It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
\underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$. After this, we can then derive
\myeqn{
    \sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out \nn
}
where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
